---
[comment]: <> (title "Weissensee results exploration 2014-2016")
[comment2]: <> (author "Ger Inberg")
[comment3]: <> (date "20-01-2016")
output: html_document
---

# Weissensee results exploration 2014-2016

## Introduction
One of my hobbies is ice-skating and in 2015 I have done the "Alternatieve Elfstedentocht" for the first time. It is an ice skating event held in Austria, over a distance of 200km ([video](https://www.youtube.com/watch?v=FhdifsMYr8s)). In a period of about 10 days multiple events are held, for professionals as well as for amateurs. 

Most participants are from the Netherlands, they are so unfortunate that there is not enough ice on their lakes to organise such an event. Because I was interested in how many people finished the 200km and what the average finish time was, I have done some data analysis.
First I have scraped the data from the Weissensee website to a [csv file](data/weissensee_results.csv), with a [python script](weissensee_scraper.py). Notice: these results are only from the amateur 200km races that are held in the years mentioned above.

[comment4]: <> (<iframe width="560" height="315" src="http://youtube.com/embed/FhdifsMYr8s"; frameborder="0" allowfullscreen></iframe>)

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that I need.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.

library(ggplot2)
library(plyr)
```

## Dataset info
```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data, combine the two datasets and print some basics
weissensee <-read.csv("data/weissensee_results.csv", header=TRUE, sep = ";") 
str(weissensee)
# add year variable for ease of use
weissensee$Year = as.numeric(format(as.POSIXct(weissensee$Date, format='%d-%m-%Y'), format="%Y"))
```

[comment5]: <> ( Univariate Plots Section)

### Finish times (200km)

The event is held at a lake where the participants skate some amount of laps to complete the 200km distance. Mostly a lap is between 10 and 25km, depending on the ice conditions. A larger lap size is supposed to lead to a faster finish time, because bigger groups can form, so there is less air resistance. Also corners slow down. The plot above shows the finish times of all the events. It follows a normal distribution quite well.

```{r echo=FALSE, message=FALSE, warning=FALSE, FinishTimes}
#finish time 200km
weissensee200km <- weissensee[as.numeric(as.matrix(weissensee$Distance)) == 200.0, ]
weissensee200km$TimeFormatted  <- as.POSIXct(strptime(weissensee200km$Time, format="%H:%M:%S"))
weissensee200km$Year = as.numeric(format(as.POSIXct(weissensee200km$Date, format='%d-%m-%Y'), format="%Y"))
ggplot(data = weissensee200km, aes(x=TimeFormatted)) + geom_histogram() + 
  ggtitle("Distribution of finish times (200km)") + xlab('Time (hours)') + 
  ylab('count')
#summary(weissensee200km$Time)
```

### Participants count

The histogram displays the number of participants per event.

```{r echo=FALSE, message=FALSE, warning=FALSE, ParticipantsCount}
# plot participate count in events
df <- count(weissensee, "Date")
#change order of levels in order to plot dates in right order
df$Date2 <- as.POSIXct(df$Date, format='%d-%m-%Y')
df$Date <- factor(df$Date, levels = df$Date[order(df$Date2)])
ggplot(data = df) + geom_histogram(aes(x=Date, y=freq), stat="identity") + 
scale_y_continuous(breaks = seq(0, 1200, 100)) + 
ggtitle("Number of participants per event") + xlab('Date') + ylab('Count')
```

### Participants country of residence

Well, I guess the histogram says it all, you can call this a Dutch event!

```{r echo=FALSE, message=FALSE, warning=FALSE, ParticipantsResidence}
#country
ggplot(data = weissensee, aes(x=Country)) + geom_histogram() + 
  scale_y_log10() + ggtitle("Country of residence of participants") + 
  xlab('Country') + ylab('count (log scale)')
summary(weissensee$Country) 
```

### Completed distance

Not all people can finish the 200km distance, because of various reasons: crashes, bad material, bad condition, etc. Fortunately, one can skate a shorter distance: simply skate less laps. The histogram shows that most people can finish the 200km, 150km is the next best.

```{r echo=FALSE, message=FALSE, warning=FALSE, CoveredDistance}
#Distance
ggplot(data = weissensee, aes(x=as.factor(Distance))) + geom_histogram() +
  scale_y_sqrt() + ggtitle("Distribution of completed distance") + 
  xlab('Distance') + ylab('count (sqrt scale)') + 
  scale_x_discrete(breaks = seq(100, 200, 10))
summary(weissensee$Distance)
```

[comment6]: <> ( Bivariate Plots Section)

### Finish times (200km) per year 

Distribution of the finish times per year. The black box is the interquartile range. Since 2016 is not yet complete, it has less points.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
# finish times per year
ggplot(data = weissensee200km, aes(x=factor(Year), y=TimeFormatted)) + 
  geom_jitter(alpha=0.1) + geom_boxplot(alpha=0.1) + 
  ggtitle("Finish time (200km) per year") + xlab('Year') + ylab('Time (hours)')
```

[comment7]: <> ( Multivariate Plots Section)
## Weissensee triple 2015

Some participants just can't get enough of ice-skating and are participating multiple times in the event. The event is held 3 times in a period of 10 days and I want to know how many people finished 3 times in the full distance and if there is a relation between their finish times.

```{r echo=FALSE, message=FALSE, warning=FALSE, Triple}
weissensee_triple <- c()
weissensee2015 <- weissensee[weissensee$Year == 2015, ]
weissensee2015_200k <- weissensee2015[as.numeric(as.matrix(weissensee2015$Distance)) == 200.0, ]
names <- unique(weissensee2015_200k$Name)
for(i in 1:length(names)){
  name = names[i]
  temp <- weissensee2015_200k[weissensee2015_200k$Name == name, ]
  if(nrow(temp) == 3){
    weissensee_triple <- rbind(weissensee_triple, temp)
  }
}
#length(unique(weissensee_triple$Name))
ggplot(data = weissensee_triple, aes(x=factor(EventId), y=Time, color=Name)) + 
  geom_point() + geom_line(aes(group=Name)) + scale_x_discrete(labels=c("23-1-15","27-1-15","30-1-15")) + 
  xlab('Date') + ylab('Finish time')
```

So, there were 6 brave man that finished 3 times! Unfortunately, it is hard to see a trend in finish times for these individuals. Some had their fastest time in the first event, some in the second and only one in the last event. Of course, the finish time depends on a lot of factors, including the individuals capabilities, weather conditions, etc.

## Ideas for future work
* relate the finish times to weather conditions (wind, snow conditions, temperature)
* compare the year 2015 to results of other years and look for trends

