---
title: "Weissensee 2015 results exploration"
author: "Ger Inberg"
date: "20-01-2016"
output: html_document
---

## Introduction
One of my hobbies is ice-skating and last year I have done the "Weissensee" for the first time. It is an ice skating event held in Austria, over a distance of 200km. The most participants are from the Netherlands. Unfortunately we don't have enough ice on our lakes, to organize such an event. Because I was interested in how many people finished the 200km and what the average finish time was, I have done some data analysis.
First I have scraped the data from the Weissensee website to a csv file, with a [python script](weissensee_scraper.py).

<iframe width="560" height="315" src="http://youtube.com/embed/FhdifsMYr8s"; frameborder="0" allowfullscreen></iframe>

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that I need.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(GGally)
```

## Dataset info
```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data, combine the two datasets and print some basics
weissensee <-read.csv("data/weissensee_results.csv", header=TRUE, sep = ";") 
str(weissensee)
```

# Univariate Plots Section
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
#plot histograms of interesting features
#finish time
weissensee$Time <- as.POSIXct(strptime(weissensee$Time, format="%H:%M:%S"))
ggplot(data = weissensee[weissensee$Laps==16,], aes(x=Time)) + geom_histogram() + 
  ggtitle("Distribution of finish times")
#summary(weissensee$Time)

#country
ggplot(data = weissensee, aes(x=Country)) + geom_histogram() + 
  scale_y_log10() + ggtitle("Country of residence of participants")
summary(weissensee$Country) 

#Laps
ggplot(data = weissensee, aes(x=Distance)) + geom_histogram() +
  scale_y_sqrt() + ggtitle("Distribution of covered distance")
summary(weissensee$Distance)
```